---
title: "Statistics for sediment toxicity"
author: "Dave Costello"
date: "May 2019"
output: html_document
---

Code for analysis presented in: 
Costello et al. 2019 Hitting reset on sediment toxicity: Sediment homogenization alters the toxicity of metal-amended sediments. *Environmental Toxicology and Chemistry.*

##Functions
```{r Functions, include=F}
#Function to calculate EC values for growth with 95% confidence intervals
ECXX=function(model,percent){
  ECname = paste("EC",percent)
  dose=names(model$dataClasses)
  EC = (100-percent)/100 
  ref = model$m$getPar()[1]
  min.x = 1
  max.x = 10000 #Might need to change the range
  range = seq(min.x,max.x,by=1)
  new.range = data.frame(frame=seq(min.x,max.x,by=1))
  names(new.range)[[1]]=dose
  bf = predict(model,newdata=new.range)
  locEC=which.min(abs(bf-ref*EC))
  se = sqrt(apply(attr(predict(model, newdata=new.range),"gradient"),1,function(x) sum(vcov(model)*outer(x,x))))
  LCI=bf+se*qnorm(0.025)
  locL=which.min(abs(LCI-ref*EC))
  UCI=bf+se*qnorm(0.975)
  locU=which.min(abs(UCI-ref*EC))
  result=c(range[locEC],range[locL],range[locU])
  result = signif(result,digits=3)
  names(result)[[1]]=ECname
  names(result)[[2]]="Lower-bound (2.5%)"
  names(result)[[3]]="Upper-bound (97.5%)"
  print(result)
  mat=cbind(new.range,bf,LCI,UCI)
  invisible(list(ResultTable=result,CImat=mat))
}

#Function to calculate LC values for survival with 95% confidence intervals
LCXX=function(model,percent){
  LCname = paste("LC",percent)
  dose = names(model$coef)[2]
  dose = ifelse(grepl("log",dose),sub("\\).*", "", sub(".*\\(", "", dose)) ,dose)
  EC = (100-percent)/100 
  min.x = 1
  max.x = 10000 #Might need to change the range
  range = seq(min.x,max.x,by=1)
  new.range = data.frame(frame=seq(min.x,max.x,by=1))
  names(new.range)[[1]]=dose
  bf = predict(model,newdata=new.range,type="response",se.fit=T)
  locLC=which.min(abs(bf$fit-EC))
  LCI=bf$fit+bf$se.fit*qnorm(0.025)
  locL=which.min(abs(LCI-EC))
  UCI=bf$fit+bf$se.fit*qnorm(0.975)
  locU=which.min(abs(UCI-EC))
  result=c(range[locLC],range[locL],range[locU])
  result = signif(result,digits=3)
  names(result)[[1]]=LCname
  names(result)[[2]]="Lower-bound (2.5%)"
  names(result)[[3]]="Upper-bound (97.5%)"
  print(result)
  mat=cbind(new.range,bf$fit,LCI,UCI)
  invisible(list(ResultTable=result,CImat=mat))
  	}

#Function to complete ANOVA-like inference from non-linear least-squares regression
ANOVA_F=function(mod,y){
  ypred=predict(mod)
  y.no.na = y[!is.na(y)]
  dfnum = length(mod$m$getPar())-1
  dfden = length(y.no.na)-length(mod$m$getPar())
  SSerr=sum((ypred-y.no.na)^2)
	SSreg=sum((ypred-mean(y.no.na))^2)
	SStot = SSerr + SSreg
  Frat = (SSreg/dfnum)/(SSerr/dfden)
  probF = pf(Frat,dfnum,dfden,lower.tail=F)
  Rsquare = 1-(SSerr/SStot)
  result=c(Frat,probF,Rsquare)
  names(result)[[1]]="F ratio"
	names(result)[[2]]="P-value"
  names(result)[[3]]="R-squared"
  return(result)
	}
```

#COPPER ANALYSES

##Import data
```{r Cu data import}
CuTox<-read.csv(file="Cu_Ha.csv")
summary(CuTox)
CuTox$Nom_Cu<-factor(CuTox$Nom_Cu)
CuTox$Dead<-10-CuTox$Count

OcoeeTox<-subset(CuTox,CuTox$Sediment=="Ocoee")
DowTox<-subset(CuTox,CuTox$Sediment=="Dow")
OcoeeTox$Nom_Cu<-OcoeeTox$Nom_Cu[,drop=T]
DowTox$Nom_Cu<-DowTox$Nom_Cu[,drop=T]
DowToxF<-subset(DowTox,DowTox$Treat=="F")
DowToxS<-subset(DowTox,DowTox$Treat=="S")
OcoeeToxF<-subset(OcoeeTox,OcoeeTox$Treat=="F")
OcoeeToxS<-subset(OcoeeTox,OcoeeTox$Treat=="S")
```

###Dow flume-aged - EC10 and LC50
```{r Dow flume-aged}
plot(Count/10~(TOT_Cu),data=DowToxF,log="x")
DFsurv <- glm(cbind(Count,Dead)~log10(TOT_Cu),data=DowToxF,family=binomial)
summary(DFsurv)
shapiro.test(resid(DFsurv)) #OK
#p < 0.001
1-DFsurv$deviance/DFsurv$null.deviance
#R2 = 0.55

DFs = LCXX(DFsurv,50) #LC50 = 435 mg/kg
for(i in 1:3){lines(DFs$CImat[,1],(DFs$CImat[,1+i]))}

plot(RGR~TOT_Cu,data=DowToxF,log="x")
DFg <- nls(RGR~SSlogis(log10(TOT_Cu),Asym,xmid,scal),data=DowToxF)
summary(DFg)
shapiro.test(resid(DFg)) #OK
ANOVA_F(DFg,DowToxF$RGR) # p < 0.001, r2 = 0.88

DF = ECXX(DFg,10) #EC10 = 57
for(i in 1:3){lines(DF$CImat[,1],DF$CImat[,1+i])}
```

###Dow spiked - EC10 and LC50
```{r Dow spiked}
plot(Count/10~TOT_Cu,data=DowToxS,log="x")
DSsurv <- glm(cbind(Count,Dead)~log10(TOT_Cu),data=DowToxS,family=binomial)
summary(DSsurv)
shapiro.test(resid(DSsurv)) #0.003
1-DSsurv$deviance/DSsurv$null.deviance
#p < 0.001, r2 = 0.74

DSs = LCXX(DSsurv,50) #LC50 395 
for(i in 1:3){lines(DSs$CImat[,1],DSs$CImat[,i+1])}

plot(RGR~TOT_Cu,data=DowToxS,log="x")
DSg <- nls(RGR~SSlogis(log10(TOT_Cu),Asym,xmid,scal),data=DowToxS)
summary(DSg)
shapiro.test(resid(DSg)) #OK
ANOVA_F(DSg,DowToxS$RGR) #p < 0.001, r2 = 0.93

DS = ECXX(DSg,10) #EC10 = 29
for(i in 1:3){lines(DS$CImat[,1],DS$CImat[,1+i])}
```


###Ocoee flume-aged - EC10 and LC50
```{r Dow flume-aged}
plot(Count/10~TOT_Cu,data=OcoeeToxF,log="x")
OFsurv <- glm(cbind(Count,Dead)~log10(TOT_Cu),data=OcoeeToxF,binomial)
summary(OFsurv)
shapiro.test(resid(OFsurv)) #OK
1-OFsurv$deviance/OFsurv$null.deviance
 #p < 0.001, r2= 0.59

OFs = LCXX(OFsurv,50) #LC50 = 2520
for(i in 1:3){lines(OFs$CImat[,1],OFs$CImat[,1+i])}

plot(RGR~TOT_Cu,data=OcoeeToxF,log="x")
OFg <- nls(RGR~SSlogis(log10(TOT_Cu),Asym,xmid,scal),data=OcoeeToxF)
summary(OFg)
shapiro.test(resid(OFg)) #OK
ANOVA_F(OFg,OcoeeToxF$RGR) #p < 0.001, r2 = 0.91

OF = ECXX(OFg,10) #EC10 = 217
for(i in 1:3){lines(OF$CImat[,1],OF$CImat[,i+1])}
```

###Ocoee spiked - EC10 and LC50
```{r Dow spiked}
plot(Count/10~TOT_Cu,data=OcoeeToxS,log="x")
OSsurv <- glm(cbind(Count,Dead)~log10(TOT_Cu),data=OcoeeToxS,binomial)
shapiro.test(resid(OSsurv)) #OK
summary(OSsurv)
1-OSsurv$deviance/OSsurv$null.deviance
#p < 0.001, r2 = 0.54

OSs = LCXX(OSsurv,50) #LC50 = 1590
for(i in 1:3){lines(OSs$CImat[,1],OSs$CImat[,1+i])}

plot(RGR~TOT_Cu,data=OcoeeToxS,log="x")
OSg <- nls(RGR~SSlogis(log10(TOT_Cu),Asym,xmid,scal),data=OcoeeToxS)
shapiro.test(resid(OSg)) #OK
summary(OSg)
ANOVA_F(OSg,OcoeeToxS$RGR) #p < 0.001, r2 = 0.89

OS = ECXX(OSg,10) #EC10 = 238
for(i in 1:3){lines(OS$CImat[,1],OS$CImat[,i+1])}
```


###Bioavailable Cu and growth response
```{r Cu bioavail}
#Dow
plot(RGR~SEMAVSfOC2,data=DowToxS,ylim=c(0,0.12),col="orangered3",las=1,log="x",xlim=c(1,1000))
points(RGR~SEMAVSfOC2,data=DowToxF,col="royalblue3",pch=15)

#Dow F alone
DFsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=DowToxF)
shapiro.test(resid(DFsao))
summary(DFsao)
ANOVA_F(DFsao,DowToxF$RGR)
DF.EC <- ECXX(DFsao,10)

#Dow S alone
DSsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=DowToxS)
shapiro.test(resid(DSsao))
summary(DSsao)
ANOVA_F(DSsao,DowToxS$RGR)
DS.EC <-ECXX(DSsao,10)

#Combined model
Dsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=DowTox)
shapiro.test(resid(Dsao))
summary(Dsao)
ANOVA_F(Dsao,DowTox$RGR)
D.EC <- ECXX(Dsao,10)

#Ocoee
plot(RGR~SEMAVSfOC2,data=OcoeeToxS,ylim=c(0,0.12),col="orangered3",las=1,log="x",xlim=c(1,10000))
points(RGR~SEMAVSfOC2,data=OcoeeToxF,col="royalblue3",pch=15)

#Ocoee F alone
OFsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=OcoeeToxF)
summary(OFsao)
shapiro.test(resid(OFsao))
ANOVA_F(OFsao,OcoeeToxF$RGR)
OF.EC <- ECXX(OFsao,10)

#Ocoee S alone
OSsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=OcoeeToxS)
summary(OSsao)
shapiro.test(resid(OSsao))
ANOVA_F(OSsao,OcoeeToxS$RGR)
OS.EC <- ECXX(OSsao,10)

#Combined model
Osao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=OcoeeTox)
shapiro.test(resid(Osao))
summary(Osao)
ANOVA_F(Osao,OcoeeTox$RGR)
O.EC <- ECXX(Osao,10)
```

#NICKEL ANALYSES

##Import data
```{r Ni data import}
NiTox<-read.csv(file="Ni_Ha.csv")
summary(NiTox)
NiTox$Nom_Ni<-factor(NiTox$Nom_Ni)
NiTox$Dead<-10-NiTox$Count

BurnTox<-subset(NiTox,NiTox$Sediment=="Burn")
RaisinTox<-subset(NiTox,NiTox$Sediment=="Raisin")
BurnTox$Nom_Ni<-BurnTox$Nom_Ni[,drop=T]
RaisinTox$Nom_Ni<-RaisinTox$Nom_Ni[,drop=T]
BurnToxF<-subset(BurnTox,BurnTox$Treat=="F")
BurnToxS<-subset(BurnTox,BurnTox$Treat=="S")
RaisinToxF<-subset(RaisinTox,RaisinTox$Treat=="F")
RaisinToxS<-subset(RaisinTox,RaisinTox$Treat=="S")
```

###Burntwood flume-aged - EC10 and LC50
```{r Burntwood flume-aged}
plot(Count/10~TOT_Ni,log="x",data=BurnToxF)
BFsurv = glm(cbind(Count,Dead)~log10(TOT_Ni),data=BurnToxF,binomial)
summary(BFsurv)
shapiro.test(resid(BFsurv)) #0.02
1-BFsurv$deviance/BFsurv$null.deviance
#p = 0.79, r2 < 0.01
#No significant mortality in any treatments

plot(RGR~TOT_Ni,log="x",data=BurnToxF)
BFg=nls(RGR~SSlogis(log10(TOT_Ni),Asym,xmid,scal),data=BurnToxF)
summary(BFg)
shapiro.test(resid(BFg)) #OK
ANOVA_F(BFg,BurnToxF$RGR) #p = 0.002, r2 = 0.44

BF = ECXX(BFg,10) #EC10 = 951 mg/kg
for(i in 1:3){lines(BF$CImat[,1],BF$CImat[,(1+i)])} 
```

###Burntwood spiked - EC10 and LC50
```{r Burntwood spiked}
plot(Count/10~TOT_Ni,log="x",data=BurnToxS)
BSsurv = glm(cbind(Count,Dead)~log10(TOT_Ni),data=BurnToxS,binomial)
summary(BSsurv)
shapiro.test(resid(BSsurv)) #OK
1-BSsurv$deviance/BSsurv$null.deviance
#p < 0.001, r2 < 0.38
#No significant mortality in any treatments

BSs = LCXX(BSsurv,50) #LC50 = 1720
for(i in 1:3){lines(BSs$CImat[,1],BSs$CImat[,1+i])}

plot(RGR~TOT_Ni,log="x",data=BurnToxS)
BSg=nls(RGR~SSlogis(log10(TOT_Ni),Asym,xmid,scal),data=BurnToxS)
summary(BSg)
shapiro.test(resid(BSg)) #OK
ANOVA_F(BSg,BurnToxS$RGR) #p < 0.001, r2 = 0.64

BS = ECXX(BSg,10) #EC10 = 1010 mg/kg
for(i in 1:3){lines(BS$CImat[,1],BS$CImat[,(1+i)])} 
```

###Raisin flume-aged - EC10 and LC50
```{r Raisin flume-aged}
plot(Count/10~TOT_Ni,log="x",data=RaisinToxF)
RFsurv = glm(cbind(Count,Dead)~log10(TOT_Ni),data=RaisinToxF,binomial)
summary(RFsurv)
shapiro.test(resid(RFsurv)) #OK
1-RFsurv$deviance/RFsurv$null.deviance
#p < 0.001, r2 = 0.54

RFs = LCXX(RFsurv,50) #LC50 = 504
for(i in 1:3){lines(RFs$CImat[,1],RFs$CImat[,1+i])}

plot(RGR~TOT_Ni,log="x",data=RaisinToxF)
RFg=nls(RGR~SSlogis(log10(TOT_Ni),Asym,xmid,scal),data=RaisinToxF)
summary(RFg)
shapiro.test(resid(RFg)) #OK
ANOVA_F(RFg,RaisinToxF$RGR) #p < 0.001, r2 = 0.85

RF = ECXX(RFg,10) #EC10 = 242 mg/kg
for(i in 1:3){lines(RF$CImat[,1],RF$CImat[,(1+i)])} 
```

###Raisin spiked - EC10 and LC50
```{r Raisin spiked}
plot(Count/10~TOT_Ni,log="x",data=RaisinToxS)
RSsurv = glm(cbind(Count,Dead)~log10(TOT_Ni),data=RaisinToxS,binomial)
summary(RSsurv)
shapiro.test(resid(RSsurv)) #OK
1-RSsurv$deviance/RSsurv$null.deviance
#p < 0.001, r2 = 0.27

RSs = LCXX(RSsurv,50) #LC50 = 803
for(i in 1:3){lines(RSs$CImat[,1],RSs$CImat[,1+i])}

plot(RGR~TOT_Ni,log="x",data=RaisinToxS)
RSg=nls(RGR~SSlogis(log10(TOT_Ni),Asym,xmid,scal),data=RaisinToxS)
summary(RSg)
shapiro.test(resid(RSg)) #0.02
ANOVA_F(RSg,RaisinToxS$RGR) #p < 0.001, r2 = 0.87

RS = ECXX(RSg,10) #EC10 = 499 mg/kg
for(i in 1:3){lines(RS$CImat[,1],RS$CImat[,(1+i)])} 
```

###Ni tissue response
```{r Ni tissue}
#Remove control treatment for regression models
BurnTox.no0 <- subset(BurnTox,Nom_Ni!="0")
RaisinTox.no0 <- subset(RaisinTox,Nom_Ni!="0")

#Burntwood
plot(Tissue_Ni~TOT_Ni,data=BurnTox,col=as.numeric(Treat),log="xy")
Btiss = lm(log10(Tissue_Ni)~log10(TOT_Ni)*Treat,data=BurnTox.no0)
summary.aov(Btiss) #Treatment effect p = 0.95
shapiro.test(resid(Btiss))

  #Simplified model without sediment prep. treatment
Btiss1 <- lm(log10(Tissue_Ni)~log10(TOT_Ni),data=BurnTox.no0) 
summary(Btiss1)
abline(Btiss1)

#Raisin
plot(Tissue_Ni~TOT_Ni,data=RaisinTox,col=as.numeric(Treat),log="xy")
Rtiss = lm(log10(Tissue_Ni)~log10(TOT_Ni)*Treat,data=RaisinTox.no0)
summary.aov(Rtiss) #Treatment effect p = 0.72
shapiro.test(resid(Rtiss)) #OK

  #Simplified model without sediment prep. treatment
Rtiss1 <- lm(log10(Tissue_Ni)~log10(TOT_Ni),data=RaisinTox.no0)
summary(Rtiss1)
abline(Rtiss1)
```


###Bioavailable Ni and growth response
```{r Ni bioavail}
#Burntwood
plot(RGR~SEMAVSfOC2,data=BurnToxS,ylim=c(0,0.12),col="orangered3",las=1,log="x",xlim=c(1,1000))
points(RGR~SEMAVSfOC2,data=BurnToxF,col="royalblue3",pch=15)

#Burntwood F alone
BFsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=BurnToxF)
shapiro.test(resid(BFsao))
summary(BFsao)
ANOVA_F(BFsao,BurnToxF$RGR)
BF.EC <- ECXX(BFsao,10)

#Dow S alone
BSsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=BurnToxS)
shapiro.test(resid(BSsao))
summary(BSsao)
ANOVA_F(BSsao,BurnToxS$RGR)
BS.EC <-ECXX(BSsao,10)

#Combined model
Bsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=BurnTox,control = list(maxiter = 500))
shapiro.test(resid(Bsao))
summary(Bsao)
ANOVA_F(Bsao,BurnTox$RGR)
B.EC <- ECXX(Bsao,10)

#Raisin
plot(RGR~SEMAVSfOC2,data=RaisinToxS,ylim=c(0,0.12),col="orangered3",las=1,log="x",xlim=c(1,1000))
points(RGR~SEMAVSfOC2,data=RaisinToxF,col="royalblue3",pch=15)

#Combined Raisin model
Rsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=RaisinTox)
shapiro.test(resid(Rsao))
summary(Rsao)
ANOVA_F(Rsao,RaisinTox$RGR)
R.EC <- ECXX(Rsao,10)
```

#ZINC ANALYSES

##Import data
```{r Zn data import}
ZnTox<-read.csv(file="Zn_Ha.csv")
summary(ZnTox)
ZnTox$Nom_Zn<-factor(ZnTox$Nom_Zn)

ZnToxF<-subset(ZnTox,ZnTox$Treat=="F")
ZnToxS<-subset(ZnTox,ZnTox$Treat=="S")
```


###Raisin (zinc) flume-aged - EC10 and LC50
```{r Raisin Zn flume-aged}
plot(Count/10~TOT_Zn,log="x",data=ZnToxF)
ZFsurv = glm(cbind(Count,Dead)~log10(TOT_Zn),data=ZnToxF,binomial)
summary(ZFsurv)
shapiro.test(resid(ZFsurv)) #OK
1-ZFsurv$deviance/ZFsurv$null.deviance
#p = 0.11, r2 = 0.03
  #Logistic model for survival is not significant

plot(RGR~TOT_Zn,log="x",data=ZnToxF)
ZFg=nls(RGR~SSlogis(log10(TOT_Zn),Asym,xmid,scal),data=ZnToxF)
summary(ZFg)
  #Cannot calculate logistic model, no reduction in growth rate
```

###Raisin (zinc) spiked - EC10 and LC50
```{r Raisin Zn spiked}
plot(Count/10~TOT_Zn,log="x",data=ZnToxS)
ZSsurv = glm(cbind(Count,Dead)~log10(TOT_Zn),data=ZnToxS,binomial)
summary(ZSsurv)
shapiro.test(resid(ZSsurv)) #OK
1-ZSsurv$deviance/ZSsurv$null.deviance
#p = 0.22, r2 = 0.03
  #Logistic model for survival is not significant

plot(RGR~TOT_Zn,log="x",data=ZnToxS)
ZSg=nls(RGR~SSlogis(log10(TOT_Zn),Asym,xmid,scal),data=ZnToxS)
summary(ZSg)
shapiro.test(resid(ZSg)) #OK
ANOVA_F(ZSg,ZnToxS$RGR) #p = 0.01, r2 = 0.30

ZS = ECXX(ZSg,10) #EC10 = 4260 mg/kg
for(i in 1:3){lines(ZS$CImat[,1],ZS$CImat[,(1+i)])} 
```

###Zn tissue response
```{r Zn tissue}
plot(Tissue_Zn~TOT_Zn,data=ZnTox,log="xy",col=c("royalblue3","orangered3")[Treat])
Ztiss = lm(log10(Tissue_Zn)~log10(TOT_Zn)*Treat,data=ZnTox)
summary.aov(Ztiss)
shapiro.test(resid(Ztiss))
abline(Ztiss)
plot(Ztiss,which=1) #Not linear

Ztiss2 <- aov(log10(Tissue_Zn)~as.factor(Nom_Zn)*Treat,data=ZnTox)
shapiro.test(resid(Ztiss2))
summary(Ztiss2)
TukeyHSD(Ztiss2)
boxplot(log10(Tissue_Zn)~Treat,data=ZnTox)
```


###Bioavailable Ni and growth response
```{r Ni bioavail}
plot(RGR~SEMAVSfOC2,data=ZnTox,col=c("royalblue3","orangered3")[Treat],log="x")
Zsao <- nls(RGR~SSlogis(log10(SEMAVSfOC2),Asym,xmid,scal),data=ZnTox)
shapiro.test(resid(Zsao)) #OK p = 0.02
summary(Zsao)
ANOVA_F(Zsao,ZnTox$RGR) #p = 0.31
  #Logistic model for growth and bioavailable Zn is not significant
```
